---
title: "ImageAnalysis"
author: "Drew Steele"
date: "2024-03-28"
output: html_document
---

```{r}
library(spatstat)
library(ggplot2)
library(gridExtra)
library(EBImage)
```

### defining variable names 

```{r}
inputDir <- "data/1344_48h_Repeat Oct 2020"
inputName <- "1344 60min PAA 48hrs 2"
outputName <- "null"
```

### setting segmentation variables 

```{r}
saveOut <- "n"

threshVal <- 0.02
gblurVal <- 1.5 
#cropRegion <- list(r = 1:689, c = 1:512)
cropRegion <- "NA"
#offsetRegion <- list(r = 940:950, c = 478:488)
offsetRegion <- "NA"
redAdj <- 1
greenAdj <- 1
waterVal <- 1
comments <- "add comment here which will be saved with other segmentation values"
keep <- "y"

threshLoop <- "n"
offsetPos <- "n"
cropReq <- "n"

aspec <- 0.5  # changes aspec ration for plotontop function
conVal <- 1.5 # changes contrast setting for work flow
```

### defining functions 

plots segmentation coordinates on top of original image 

```{r}
plotontop <- function(img, live, dead, xcord, ycord, title) {
  
  x <- flip(img * conVal)
  
  par(mar = c(0, 0, 0, 0))  # Set margins to zero
  plot(1, type = "n", xlab = "", ylab = "", xlim = c(0, 1), ylim = c(0, aspec), asp =1, main = paste(title))
  
  rasterImage(x, 0, 0, 1, dim(x)[2]/dim(x)[1])
  points(x = live[,xcord] / dim(x)[1], y = live[,ycord]/ dim(x)[1], col = "blue", pch = 1, cex = 1)
  points(x = dead[,xcord] / dim(x)[1], y = dead[,ycord] / dim(x)[1], col = "red", pch = 1, cex = 1)
}
```


### load original segmentation + image and view 

```{r}
OGgreen <- read.csv(paste0(inputDir, "/", inputName, ".green.csv"))
OGred <- read.csv(paste0(inputDir, "/", inputName, ".red.csv"))

imgRaw <- readImage(paste0(inputDir, "/", inputName, ".tif"))
display(flip(imgRaw), method="browser")
display(flip(imgRaw * conVal), method="browser")
#display(imgRaw *2, method="browser")
```

### cropping image 

```{r}
if(cropReq == "y") {
  
  img <- imgRaw[cropRegion$r, cropRegion$c, 1:3]
  
  par(mfrow = c(1,1))
  plot(img)


} else {
  
  cat("no cropping required")
  
  img <- imgRaw
  
}
```


### splitting and combining colour channels 

```{r}
imgR <- getFrame(img, 1) # splitting colour channels
imgG <- getFrame(img, 2)

imgC <- normalize((imgR * redAdj) + (imgG * greenAdj), ft=c(0,1)) # creating flatted image

par(mfrow=c(3,1))
plot(imgR)
plot(imgC)
plot(imgG)

par(mar = c(6,6,6,6))  # histograms of colour intensities
par(mfrow = c(2,2))
hist(imgR, main = "red")
hist(imgG,main = "green")
hist(img,main = "OG")
hist(imgC,main = "combo")

display(imgC * conVal)
```

### offset subtraction 

```{r}
if(offsetPos == "y") {
  
  par(mfrow = c(1,2))
  plot(imgC[offsetRegion$r, offsetRegion$c]*4)

  os=mean(imgC[offsetRegion$r, offsetRegion$c])

  imgCo <- normalize(imgC - os, ft=c(0,1)) # subtracting mean intensity from full image 

  par(mfrow = c(1,2))
  plot(imgC)
  plot(imgCo)


} else {
  
  cat("no offset possible")
  
  imgCo <- normalize(imgC, ft=c(0,1))
  
}
```

### edge detection

```{r}
k = matrix(-1, nrow = 3, ncol = 3) # defining convolution kernel 
k[2, 2] = 8
k

imgE <- EBImage::filter2(gblur(imgCo,sigma=gblurVal),k) # applying edge detection and Gaussian blur 

par(mfrow = c(2,1))
plot(img * conVal)
plot(normalize(imgE))
```

### thresholding

```{r}
if(threshLoop == "y") { # running loop to find best threshold value 
  
  for (i in seq(from = 0.005, to = 0.05, by = 0.005)) {
  
    x <- thresh(imgE, w=5, h=5, offset= i)
  
    par(mfrow = c(2,1))
    display(x, method="raster")
    text(x = 20, y = 20, label = paste("offset = ", i), adj = c(0,1), col = "green", cex = 1.5)
    display(img * conVal, method="raster")
    text(x = 20, y = 20, label = "Original image", adj = c(0,1), col = "green", cex = 1.5)
  
  }
} else {
  cat("green threshold loop not run")
}

imgTh <- thresh(imgE, w=5, h=5, offset = threshVal) # applying threshold, w and h = kernel size for adaptive thresh
plot(img)
plot(imgTh)
```

### eroding + dilation remove single pixels 

```{r}
#kern <- makeBrush(3, shape = "box")
#logo_erode= erode(imgTh, kern)
#logo_dilate = dilate(logo_erode, kern)

#plot(logo_dilate)
#plot(imgTh)
```

watershed algorithm 

```{r}
nmask <- watershed(EBImage::distmap(imgTh), tolerance = waterVal, ext = 1)

par(mfrow = c(2,1))
plot(colorLabels(nmask), all = T)
plot(img * conVal)
```

extracting intensity vals and cell coordinates - assigning live dead status 

```{r}
IntValR <- as.data.frame(computeFeatures.basic(nmask, imgR)[,1:2]) # extractign intensity info for each cell - using red image as ref
IntValG <- as.data.frame(computeFeatures.basic(nmask, imgG)[,1:2]) # using green image as ref

cellLoc <- as.data.frame(cbind(computeFeatures.moment(nmask)[,1:2], IntValR, IntValG)) # extracting cell coordinates
colnames(cellLoc) <- c("m.cx", "m.cy","redMeanInts", "redSDInts", "greenMeanInts", "greenSDInts")
cellLoc$condition <- "null"

deadSet <- which(cellLoc$redMeanInts > cellLoc$greenMeanInts) # assigning dead set 
liveSet <- which(cellLoc$greenMeanInts > cellLoc$redMeanInts) # assigning live set 

cellLoc$condition[deadSet] <- "0" 
cellLoc$condition[liveSet] <- "1" 
```

### plotting original and new segmentation 

```{r}
par(mfrow = c(1,1))

ggplot(cellLoc, mapping = aes(x = redMeanInts, y = greenMeanInts, colour = condition)) + # plotting red and green intensity for each cell 
  geom_point() +
  xlim(c(0,1)) +
  ylim(c(0,1)) +
  ggtitle("live dead distribution") + 
  theme(aspect.ratio=1)

plotontop(imgRaw, OGgreen, OGred, "XM", "YM", "original segmentation") # comparing to original 

plotontop(img, cellLoc[cellLoc$condition == "1" ,], 
          cellLoc[cellLoc$condition == "0" ,], "m.cx", "m.cy", "new segmentation") # examining new segmentation 
```

### creating data frame for segmentation values 

```{r}
varNames <- c("threshVal", "gblurVal", "offsetRegion", "cropRegion", "sumR", "sumG", 
            "meanG", "meanR", "redAdj", "greenAdj", "deadCount", "liveCount", "comments", "keep")

varVal <- c(threshVal, gblurVal, paste(offsetRegion, collapse = " "), 
            paste(cropRegion, collapse = " "), sum(imgR), sum(imgG), 
            mean(cellLoc$greenMeanInts), mean(cellLoc$redMeanInts), redAdj, greenAdj, 
            sum(cellLoc$condition == "0"), sum(cellLoc$condition == "1"), comments, keep)

segVals <- data.frame(varNames = varNames, varVal = varVal)
```

### saving output 

```{r}
if(saveOut == "y") {
  
  write.csv(cellLoc, paste0("results_good/segmentation2/", outputName, ".csv"))
  write.table(segVals, paste0("results_good/segmentation2/", outputName, "_values.txt"), sep = "\t")
  
} else {
  cat("not saved")
}
```
