---
title: "SpatialAnalysis"
author: "Drew Steele"
date: "2024-03-09"
output: html_document
---

### script for feature extraction and spatial analysis

```{r}
aspec <- 0.7

plotontop <- function(img, cells, red, xcord, ycord, title) {
  
  x <- flip(img * 2)
  
  par(mar = c(0, 0, 0, 0))  # Set margins to zero
  plot(1, type = "n", xlab = "", ylab = "", xlim = c(0, 1), ylim = c(0, aspec), asp =1, main = paste(title))
  
  rasterImage(x, 0, 0, 1, dim(x)[2]/dim(x)[1])
  points(x = cells[,xcord] / dim(x)[1], y = cells[,ycord]/ dim(x)[1], col = "blue", pch = 1, cex = 1)
  points(x = red[,xcord] / dim(x)[1], y = red[,ycord] / dim(x)[1], col = "red", pch = 1, cex = 1)
}
```

```{r}
library(spatstat)
library(deldir)
```

### defining variables 

```{r}
imgDir <- "data/1344_48h_Repeat Oct 2020"
imgName <- "1344 40min PAA 48hrs 3"

inputName <- "1_48_PAA_60_2"

len <- 1024 # setting image size for cropped images 
hei <- 512
```

### loading data and manipulating 

```{r}

dat <- read.csv(paste0("results_good/segmentation/", inputName, ".csv"))
fullcells <- dat[,c("m.cx", "m.cy", "condition")]
fullcells$condition[fullcells$condition == "null"] <- "0"
fullcells$condition <- as.factor(fullcells$condition)
fullcells$id <- rownames(fullcells)
str(fullcells)

living <- fullcells[fullcells$condition == "1" ,]
dead <- fullcells[fullcells$condition == "0" ,]

summary(fullcells)

```

### viewing segmentation output 

```{r}
img <- readImage(paste0(imgDir, "/", imgName, ".tif"))

plotontop(img, fullcells[fullcells$condition == "1" ,], 
          fullcells[fullcells$condition == "0" ,], "m.cx", "m.cy", "new segmentation") # examining segmentation 
```

###  creating point pattern object 

```{r}

pat <- ppp(fullcells[,1], fullcells[,2], c(0,len), c(0,hei)) # creating point pattern object 
marks(pat) <- fullcells[,3] # adding info on live dead - any extra info refereed to as mark in spatstat

patDead <- ppp(dead[,1], dead[,2], c(0,len), c(0,hei)) # creating point pattern object 

patLive <- ppp(living[,1], living[,2], c(0,len), c(0,hei)) # creating point pattern object 

```

### exploring point pattern objects

```{r}

plot(pat)
anyDuplicated(pat) # checking for duplicated points 
summary(pat)

plot(patDead)
anyDuplicated(patDead) 
summary(patDead)

plot(patLive)
anyDuplicated(patLive)
summary(patLive)

```

### Spatial randomness test 

```{r}
QAll <- quadratcount(pat, nx = 20, ny = 10)
plot(pat)
plot(QAll, add = TRUE, cex = 1)

quadrat.test(QAll, alternative="two.sided")
quadrat.test(QAll, alternative="regular")
quadrat.test(QAll, alternative="clustered")

QDead <- quadratcount(patDead, nx = 20, ny = 10)
plot(patDead)
plot(QDead, add = TRUE, cex = 1)

quadrat.test(QDead, alternative="two.sided")
quadrat.test(QDead, alternative="regular")
quadrat.test(QDead, alternative="clustered")

QLive <- quadratcount(patLive, nx = 20, ny = 10)
plot(patLive)
plot(QLive, add = TRUE, cex = 1)

quadrat.test(QLive, alternative="two.sided")
quadrat.test(QLive, alternative="regular")
quadrat.test(QLive, alternative="clustered")
```

### density tests 

```{r}
denAll <- density(pat) # using default bandwidth 
attr(denAll, "sigma") # shows what bandwidth defaulted to 
plot(denAll, main = "All") #plotting intensities 

denDead <- density(patDead)
attr(denDead, "sigma")
plot(denDead, main = "Dead")

denLive <- density(patLive)
attr(denLive, "sigma")
plot(denLive, main = "Live")

densityVal <- denAll[pat] # extracting intensity values for each point 
```

### extracting distance data - to any cell 

```{r}

NNdist <- as.data.frame(nndist(pat, k = 1:6))
NNdist[,"meanDist"] <- rowMeans(NNdist[, 1:6], na.rm = TRUE)
NNdist[,"sumDist"] <- rowSums(NNdist[, 1:6], na.rm = TRUE)

```

### extracting distance data - to live / dead cell 

```{r}

NNdeadLive <- nncross(patDead, patLive, what = "dist", k = 1:6)
NNliveDead <- nncross(patLive, patDead, what = "dist", k = 1:6)
NNdeadDead <- nndist(patDead, k = 1:6)
NNliveLive <- nndist(patLive, k = 1:6)

```

### extracting how many living and how many dead neighbours in certain radius

```{r}

# defining function to use in the marksat function 
NEcounts <- function(x) {
  alive_count <- sum(x == 1)
  dead_count <- sum(x == 0)
  
  return(c(aliveN = alive_count, deadN = dead_count))
}

neighbourInf <- t(markstat(pat, NEcounts, R = 15))

```

### Voronoi tesselation transformation

```{r}

voronoi <- deldir(fullcells[,1:2], z = fullcells[,3], rw=c(0, len, 0, hei), eps=1e-09, sort=TRUE, plot=T,
                  round=TRUE,digits=6, id=fullcells[,4])

print(voronoi)

vorInf <- voronoi$summary
nbrs <- getNbrs(voronoi) # extracting neighbour information 

```

### extraactign live dead information for each cells neigbhours

```{r}

dead.live <- matrix(nrow = length(nbrs), ncol = 2)
colnames(dead.live) <- c("deadNV", "liveNV")


for (i in seq_along(nbrs)) {
  x <- as.character(nbrs[[i]])
  Ninf <- as.character(vorInf[x, "z"])
  dead <- as.numeric(sum(Ninf == "0"))
  live <- as.numeric(sum(Ninf == "1"))
  dead.live[i, "deadNV"] <- dead
  dead.live[i, "liveNV"] <- live
}

```

### creating data frame of output 

```{r}
deadDLdist <- as.data.frame(cbind(NNdeadLive, NNdeadDead))
  #colnames(deadDLdist) <- c("distLive.1", "distLive.2", "distLive.3", "distDead.1", "distDead.2", "distDead.3")
liveDLdist <- as.data.frame(cbind(NNliveLive, NNliveDead))
  #colnames(liveDLdist) <- c("distLive.1", "distLive.2", "distLive.3", "distDead.1", "distDead.2", "distDead.3")
  
DLdist <- rbind(liveDLdist, deadDLdist)
DLdist$liveMeanDist <- rowMeans(DLdist[, 1:6], na.rm = TRUE)
DLdist$deadMeanDist <- rowMeans(DLdist[, 7:12], na.rm = TRUE)

cellInf <- cbind(fullcells, densityVal, NNdist[, c("dist.1", "dist.2", "dist.3", "meanDist")], neighbourInf, 
                 DLdist[, c("liveMeanDist", "deadMeanDist")], dead.live, 
                 vorInf[, c("n.tri", "n.tside", "nbpt", "dir.area")])

cellInf <- cellInf[cellInf$nbpt == 0 ,]


deadInf <- cellInf[cellInf$condition == 0 ,]
liveInf <- cellInf[cellInf$condition == 1 ,]

```

### exploring outputs 

```{r}
#par(mfrow = c(1,2))

#for (i in c("densityVal", "dist.1", "dist.2", "dist.3", "meanDist", "sumDist", "aliveN", "deadN",
#            "distLive.1", "distLive.2", "distDead.1", "distDead.2")) {
  
#  cat("Variable:", i, "\n")
  
#  cat("Mean (Dead):", mean(deadInf[[i]]), "\n")
#  cat("Median (Dead):", median(deadInf[[i]]), "\n")

#  cat("Mean (Live):", mean(liveInf[[i]]), "\n")
#  cat("Median (Live):", median(liveInf[[i]]), "\n")

  # Plot histograms
#  hist(deadInf[[i]], main = paste("Histogram (Dead) -", i), xlab = i, col = "lightblue", border = "black")
#  hist(liveInf[[i]], main = paste("Histogram (Live) -", i), xlab = i, col = "lightgreen", border = "black")

  # Pause for user input (press Enter to continue)
#  cat("Press Enter to continue...")
#  invisible(readLines(n = 1))
#}


```

### saving output 

```{r}

#write.csv(deadInf, paste0("results_good/CellInfo/", inputName, ".Dead.Inf.csv"))
#write.csv(liveInf, paste0("results_good/CellInfo/", inputName, ".Live.Inf.csv"))
write.csv(cellInf, paste0("results_good/CellInfo/", inputName, ".All.Inf.csv"))

```

