---
title: "RandomForestQC_evaluation"
author: "Drew Steele"
date: "2024-04-06"
output: html_document
---

### evalauation of random forrest quantile classifer and data imbalance techniques 

```{r}

library(randomForestSRC)
library(caret)
library(ROSE)
library(imbalance)
library(corrplot)
library(nortest)
library(GGally)

```

### defining variables 

```{r}
dir1 <- "results_good/CellInfo/"

datName1 <- "1_48_PAA_40_2"
datName2 <- "1_48_PAA_40_3"
```

### loading data 

```{r}
dat1 <- read.csv(paste0(dir1, datName1, ".All.Inf.csv"))
dat1$condition <- as.factor(dat1$condition)

dat2 <- read.csv(paste0(dir1, datName2, ".All.Inf.csv"))
dat2$condition <- as.factor(dat2$condition)
  
datAll <- rbind(dat1,dat2)
  
str(dat1)
str(dat2)
str(datAll)
```

### correlation test 

```{r}
par(mfrow = c(1,1))

corVals <- cor(datAll[,6:17])

corrplot(corVals)
```

### selecting features

```{r}
varSelect <- c("condition", "densityVal", "liveMeanDist", "deadMeanDist", "deadNV", "liveNV", "n.tri", "dir.area")
#varSelect <- c("condition", "densityVal", "deadNV", "liveNV", "n.tri", "dir.area")
#varSelect <- c("condition", "deadNV", "liveNV", "n.tri")
#varSelect <- c("condition", "densityVal", "n.tri", "dir.area")
```

### splitting data 

```{r}
set.seed(101)


train1 <- dat1[, varSelect]
train2 <- dat2[, varSelect]

indexAll <- createDataPartition(datAll$condition, p = 0.7, list = FALSE) # taking 40% of data
trainAll <- datAll[indexAll, varSelect]
testAll <- datAll[-indexAll, varSelect]

summary(train1)
summary(train2)
summary(trainAll)
summary(testAll)
```

### creating synthetic data with SMOTE 

```{r}
set.seed(101)

newDat1 <- mwmote(train1, numInstances = 200, classAttr = "condition")
newDat2 <- mwmote(train2, numInstances = 200, classAttr = "condition")
newDatAll <- mwmote(trainAll, numInstances = 200, classAttr = "condition")

SMOTEdat1 <- rbind(newDat1, train1)
SMOTEdat2 <- rbind(newDat2, train2)
SMOTEdatAll <- rbind(newDatAll, trainAll)
```

### splitting data - over + under sampling 

```{r}
set.seed(101)

dataBalance1 <- ovun.sample(condition~., data=train1, p=0.4,
                        seed=1, method="both")$data

dataBalance2 <- ovun.sample(condition~., data=train2, p=0.4,
                        seed=1, method="both")$data

dataBalanceAll <- ovun.sample(condition~., data=trainAll, p=0.4,
                        seed=1, method="both")$data

sum(dataBalance1$condition == "0")
sum(dataBalance1$condition == "1")
```

### creating models

```{r}
set.seed(101)

mod.1.RFQ <- imbalanced(condition ~ ., train1, importance = T)
mod.2.RFQ <- imbalanced(condition ~ ., train2, importance = T)
mod.All.RFQ <- imbalanced(condition ~ ., trainAll, importance = T)

mod.1.RFQ.ovun <- imbalanced(condition ~ ., dataBalance1, importance = T)
mod.2.RFQ.ovun <- imbalanced(condition ~ ., dataBalance2, importance = T)
mod.All.RFQ.ovun <- imbalanced(condition ~ ., dataBalanceAll, importance = T)

mod.1.RFQ.SMOTE <- imbalanced(condition ~ ., SMOTEdat1, importance = T)
mod.2.RFQ.SMOTE <- imbalanced(condition ~ ., SMOTEdat2, importance = T)
mod.All.RFQ.SMOTE <- imbalanced(condition ~ ., SMOTEdatAll, importance = T)
```

### evaluating models - RFQ

```{r}
par(mar = c(3,3,3,3))

print(mod.1.RFQ)
print(predict(mod.1.RFQ, train2))
plot(mod.1.RFQ)


print(mod.2.RFQ)
print(predict(mod.2.RFQ, train1))
plot(mod.2.RFQ)


print(mod.All.RFQ)
print(predict(mod.All.RFQ, testAll))
plot(mod.All.RFQ)
```

### evaluating models - RFQ ovun smapling

```{r}
par(mar = c(3,3,3,3))

print(mod.1.RFQ.ovun)
print(predict(mod.1.RFQ.ovun, train2))
plot(mod.1.RFQ.ovun)


print(mod.2.RFQ.ovun)
print(predict(mod.2.RFQ.ovun, train1))
plot(mod.2.RFQ.ovun)


print(mod.All.RFQ.ovun)
print(predict(mod.All.RFQ.ovun, testAll))
plot(mod.All.RFQ.ovun)
```

### evaluating models - RFQ SMOTE

```{r}
par(mar = c(3,3,3,3))

print(mod.1.RFQ.SMOTE)
print(predict(mod.1.RFQ.SMOTE, train2))
plot(mod.1.RFQ.SMOTE)


print(mod.2.RFQ.SMOTE)
print(predict(mod.2.RFQ.SMOTE, train1))
plot(mod.2.RFQ.SMOTE)


print(mod.All.RFQ.SMOTE)
print(predict(mod.All.RFQ.SMOTE, testAll))
plot(mod.All.RFQ.SMOTE)
```

### calculating variable importance using subsampliong for confidence intervals - RFQ

```{r}
set.seed(101)

mod.1sub.RFQ <- subsample(mod.1.RFQ, verbose = T)
mod.2sub.RFQ <- subsample(mod.2.RFQ, verbose = T)
mod.Allsub.RFQ <- subsample(mod.All.RFQ, verbose = T)


par(mfrow = c(1,3))
plot.subsample(mod.1sub.RFQ, main = "Mod 1")
plot.subsample(mod.2sub.RFQ, main = "Mod 2")
plot.subsample(mod.Allsub.RFQ, main = "Mod All")
```

### calculating variable importance using subsampliong for confidence intervals - ovun under/overs sampling

```{r}
set.seed(101)

mod.1sub.RFQ.ovun <- subsample(mod.1.RFQ.ovun, verbose = T)
mod.2sub.RFQ.ovun <- subsample(mod.2.RFQ.ovun, verbose = T)
mod.Allsub..RFQ.ovun <- subsample(mod.All.RFQ.ovun, verbose = T)


par(mfrow = c(1,3))
plot.subsample(mod.1sub.RFQ.ovun, main = "Mod 1")
plot.subsample(mod.2sub.RFQ.ovun, main = "Mod 2")
plot.subsample(mod.Allsub..RFQ.ovun, main = "Mod All")
```


### calculating variable importance using subsampliong for confidence intervals - SMOTE

```{r}
set.seed(101)

mod.1sub.RFQ.SMOTE <- subsample(mod.1.RFQ.SMOTE, verbose = T)
mod.2sub.RFQ.SMOTE <- subsample(mod.2.RFQ.SMOTE, verbose = T)
mod.Allsub..RFQ.SMOTE <- subsample(mod.All.RFQ.SMOTE, verbose = T)


par(mfrow = c(1,3))
plot.subsample(mod.1sub.RFQ.SMOTE, main = "Mod 1")
plot.subsample(mod.2sub.RFQ.SMOTE, main = "Mod 2")
plot.subsample(mod.Allsub..RFQ.SMOTE, main = "Mod All")
```

