---
title: "cell_segment_presentation_stuff"
author: "Drew"
date: "2024-02-23"
output: html_document
---

```{r}
library(EBImage)
```

```{r}
img <- readImage("data/trial2.tif")
```

```{r}
display(img, method="browser")
```

cropping 

```{r}
#imgCr = img[512:1024, 1:512, 1:3]
imgCr <- img

imgR <- getFrame(imgCr, 1)
imgG <- getFrame(imgCr, 2)

par(mfrow = c(3,1))
plot(imgR)
plot(imgG)
plot(imgCr)


par(mfrow = c(1,2))
plot(imgG[948:975, 325:360]*4)
plot(imgR[948:975, 325:360]*4)

osR=mean(imgR[948:975, 325:360])
osG=mean(imgG[948:975, 325:360])

imgRo <- imgR - osR
imgGo <- imgG - osG

par(mfrow = c(2,2))
plot(imgR)
plot(imgRo)
plot(imgG)
plot(imgGo)
```

offset subtraction - cropped 


cropping 

```{r}
imgCr = img[512:1024, 1:512, 1:3]
imgR <- getFrame(imgCr, 1)
imgG <- getFrame(imgCr, 2)

par(mfrow = c(1,2))
plot(imgG[436:463, 325:360]*4)
plot(imgR[436:463, 325:360]*4)

osR=mean(imgR[436:463, 325:360])
osG=mean(imgG[436:463, 325:360])

imgRo <- imgR - osR
imgGo <- imgG - osG

par(mfrow = c(2,2))
plot(imgR)
plot(imgRo)
plot(imgG)
plot(imgGo)

```

edge detection

```{r}

k = matrix(-1, nrow = 3, ncol = 3)
k[2, 2] = 8
k

imgGe <- EBImage::filter2(gblur(imgGo,sigma=1.5),k)
par(mfrow = c(2,1))
plot(imgCr)
plot(normalize(imgGe))


Shoriz <- matrix(c(1, 2, 1, 0, 0, 0, -1, -2, -1), nrow = 3)
Svert <- t(Shoriz)

# get horizontal and vertical edges
imgHg <- filter2(imgGo, Shoriz)
imgVg <- filter2(imgGo, Svert)

# combine edge pixel data to get overall edge data
hdataG <- imageData(imgHg)
vdataG <- imageData(imgVg)
edataG <- sqrt(hdataG^2 + vdataG^2)

# transform edge data to image
imgGe2 <- Image(edataG)
# print(display(combine(img, imgH, imgV, imgE), method = "raster", all = T))

par(mfrow = c(1,2))
plot(img)
plot(imgGe2)

par(mfrow = c(1,1))
display(normalize(imgGe), method = "raster")
#ext(x = 20, y = 20, label = "Laplacian of Gaussian", adj = c(0,1), col = "green", cex = 1.4)
display(imgGe2, method="raster")
text(x = 20, y = 20, label = "Sobel Operator", adj = c(0,1), col = "green", cex = 1.4)
```


adaptive thresholding - Edge2 image 

```{r}

imgTh1 <- thresh(imgGo, w=5, h=5, offset=0.035)
imgTh2 <- thresh(imgGe, w=5, h=5, offset=0.035)
#imgTh2 <- img > intermodes(imgGe)
  
mean(imgGe)


par(mfrow = c(2,2))
display(imgTh1, method = "raster")
text(x = 20, y = 20, label = "Offset = 0.035", adj = c(0,1), col = "green", cex = 1.8)
display(imgTh2, method = "raster")

```

watershed

```{r}
nmask1 <- watershed(distmap(imgTh1), tolerance = 1, ext = 1)
nmask2 <- watershed(distmap(imgTh2), tolerance = 1, ext = 1)
nmask3 <- watershed(distmap(imgTh3), tolerance = 1, ext = 1)

display(colorLabels(nmask1), all=TRUE)
par(mfrow = c(2,1))
plot(colorLabels(nmask1))
plot(imgCr * 2)

display(colorLabels(nmask2), all=TRUE)
par(mfrow = c(2,1))
plot(colorLabels(nmask2))
plot(imgCr * 2)

display(colorLabels(nmask3), all=TRUE)
par(mfrow = c(2,1))
plot(colorLabels(nmask3))
plot(imgCr * 2)

# nmask

#display(colorLabels(nmaske2), all=TRUE)

par(mfrow = c(1,1))
plot(colorLabels(nmaskG))
```

```{r}
cellLocA <- as.data.frame(computeFeatures.moment(nmask3))

# write.csv(all_cells, file = "results/all_cells.csv")
```

### extractign just red cells 

making red bolder 

```{r}
imgRb <- imgRo - imgGo /1.9

par(mfrow = c(1,3))
display(imgRo, method="raster")
text(x = 20, y = 20, label = "Red spectrum", adj = c(0,1), col = "green", cex = 2)
display(imgRb*1.5, method="raster")
text(x = 20, y = 20, label = "Ratio image", adj = c(0,1), col = "green", cex = 2)
display(imgCr, method="raster")
text(x = 20, y = 20, label = "Original image", adj = c(0,1), col = "green", cex = 2)
```

edge detection - red only 

```{r}

k = matrix(-1, nrow = 3, ncol = 3)
k[2, 2] = 8
k

imgRbe <- EBImage::filter2(gblur(imgRb,sigma=1.5),k)
par(mfrow = c(2,1))
plot(imgCr)
plot(normalize(imgRbe))
```

adaptive thresh - red 

```{r}
thRb1 <- thresh(imgRbe, w=5, h=5, offset=0.04)
thRb2 <- thresh(imgRbe, w=5, h=5, offset=0.036)
thRb3 <- thresh(imgRbe, w=5, h=5, offset=0.033)

par(mfrow = c(3,1))
plot(imgRb)
plot(thRb1)
plot(imgCr *2)

par(mfrow = c(3,1))
plot(imgRb)
plot(thRb2)
plot(imgCr *2)

par(mfrow = c(3,1))
plot(imgRb)
plot(thRb3)
plot(imgCr *2)

```

watershed on red cells 

```{r}

nmaskR <- watershed(distmap(thRb1), tolerance = 1, ext = 1)

display(colorLabels(nmaskR), all=TRUE)


par(mfrow = c(1,2))
display(colorLabels(nmaskR), method="raster")
text(x = 20, y = 20, label = "Dead Cell Watershed", adj = c(0,1), col = "green", cex = 1.5)
display(imgCr, method="raster")
text(x = 20, y = 20, label = "Original Image", adj = c(0,1), col = "green", cex = 1.5)


cellLocR <- as.data.frame(computeFeatures.moment(nmaskR))

```

defining function to plot on original image 

```{r}
plotontop <- function(img, cells, red, xcord, ycord) {
  
  x <- flip(img *2)
  
  par(mar = c(0, 0, 0, 0))  # Set margins to zero
  plot(1, type = "n", xlab = "", ylab = "", xlim = c(0, 1), ylim = c(0, 1), asp =1)
  
  rasterImage(x, 0, 0, 1, dim(x)[2]/dim(x)[1])
  points(x = cells[,xcord] / dim(x)[1], y = cells[,ycord]/ dim(x)[1], col = "blue", pch = 20, cex = 0.3)
  points(x = red[,xcord] / dim(x)[1], y = red[,ycord] / dim(x)[1], col = "red", pch = 20, cex = 0.3)
  
}

plotontop2 <- function(img, cells, red, xcord, ycord) {
  
  x <- flip(img *2)
  
  par(mar = c(0, 0, 0, 0))  # Set margins to zero
  plot(1, type = "n", xlab = "", ylab = "", xlim = c(0, 1), ylim = c(0, 1), asp =1)
  
  rasterImage(x, 0, 0, 1, dim(x)[2]/dim(x)[1])
  points(x = cells[,xcord] / dim(x)[1], y = cells[,ycord]/ dim(x)[1], col = "blue", pch = 1, cex = 0.7)
  points(x = red[,xcord] / dim(x)[1], y = red[,ycord] / dim(x)[1], col = "red", pch = 1, cex = 0.7)
}

plotontop3 <- function(img, cells, xcord, ycord) {
  
  x <- flip(img*1.5)
  
  par(mar = c(0, 0, 0, 0))  # Set margins to zero
  plot(1, type = "n", xlab = "", ylab = "", xlim = c(0, 1), ylim = c(0, 1), asp =1)
  
  rasterImage(x, 0, 0, 1, dim(x)[2]/dim(x)[1])
  points(x = cells[,xcord] / dim(x)[1], y = cells[,ycord]/ dim(x)[1], col = "blue", pch = 1, cex = 0.7)
  #points(x = red[,xcord] / dim(x)[1], y = red[,ycord] / dim(x)[1], col = "red", pch = 1, cex = 0.7)
}

plotontop4 <- function(img, red, xcord, ycord) {
  
  x <- flip(img*1.5)
  
  par(mar = c(0, 0, 0, 0))  # Set margins to zero
  plot(1, type = "n", xlab = "", ylab = "", xlim = c(0, 1), ylim = c(0, 1), asp =1)
  
  rasterImage(x, 0, 0, 1, dim(x)[2]/dim(x)[1])
  #points(x = cells[,xcord] / dim(x)[1], y = cells[,ycord]/ dim(x)[1], col = "blue", pch = 1, cex = 0.7)
  points(x = red[,xcord] / dim(x)[1], y = red[,ycord] / dim(x)[1], col = "red", pch = 1, cex = 0.7)
}
```

plotting red and green extracted points on image 

```{r}
par(mfrow = c(1,1))

plotontop2(imgCr, cellLocA, cellLocR, "m.cx", "m.cy")
plotontop3(imgCr, cellLocA, "m.cx", "m.cy")
plotontop4(imgCr, cellLocR, "m.cx", "m.cy")


```

### now goign to calc neareast neighbour, if neigbour is within certain distance will delete the poitn from all cells and only keep the red cell 

```{r}
library(spatstat)
```

manipulating data

```{r}
red_c <- cellLocR[,1:2]
red_c[,3] <- "dead"

all_c <- cellLocA[,1:2]
all_c[,3] <- "all"
 
cells <- rbind(red_c, all_c)
colnames(cells) <- c("m.cx", "m.cy", "condition")
```

creating spatstat objects

```{r}

par(mar=c(2,2,2,2))

patAll <- ppp(cellLocA[,1], cellLocA[,2], c(0,513), c(0,512)) # creating point pattaern object 

anyDuplicated(patAll) # checking for duplicated poitns 

plot(patAll)

patRed <- ppp(cellLocR[,1], cellLocR[,2], c(0,513), c(0,512)) # creating point pattaern object 

anyDuplicated(patRed) # checking for duplicated poitns 

plot(patRed)

pat <- ppp(cells[,1], cells[,2], c(0,513), c(0,512)) # creating point pattaern object 

marks(pat) <- cells[,3] # adding infor on live dead - any extra info refered to as mark in spatstat

anyDuplicated(pat) # checking for duplicated poitns 

plot(pat, main = "Point Pattern Object")

```

calculating nearest neighbour and selecting those closet together 

```{r}

NN_which <- nnwhich(cells)
NN_dist <- nndist(cells)

NN_dist_All <- nndist(patAll)
NN_dist_red <- nndist(patRed)

NN_all <- as.data.frame(cbind(NN_which, NN_dist))

```

examining dist data

```{r}

min(NN_all$NN_dist)
max(NN_all$NN_dist)
hist(NN_all$NN_dist)

min(NN_dist_All)
max(NN_dist_All)
hist(NN_dist_All)

min(NN_dist_red)
max(NN_dist_red)
hist(NN_dist_red)

```


combining cell coords with distance - removing close points 

```{r}

min(NN_dist_All) # checking smallest distance between all cells

cellDists <- cbind(cells, NN_all)

cellLocDone <- as.data.frame(cellDists[cellDists$condition == "all" & cellDists$NN_dist > 2.692582 ,])

#your_data$condition <- ifelse(your_data$condition == "live" & your_data$NN_dist < 2.6, "dead", your_data$condition)

```

loading in original points 

```{r}

redOG <- read.csv("data/redPAA48.csv")
greenOG <- read.csv("data/greenPAA48.csv")

```

exaimining success 

```{r}

plotontop2(imgCr, cellLocDone, cellLocR, "m.cx", "m.cy") # examining my segmentation 

plotontop2(imgCr, greenOG, redOG, "XM", "YM") # comparing to original 

```

saving outputs 

```{r}
write.csv(cellLocDone, "results/livingCellsLoc.csv")
write.csv(cellLocR, "results/deadCellsLoc.csv")
```


### calculating neighbours in x distance 

data prep 

```{r}
live <- livingcells[,1:2]
live[,3] <- 1
colnames(live) <- c("x", "y", "condition")

dead <- red_cells[,2:3]
dead[,3] <- 0
colnames(dead) <- c("x", "y", "condition")

fullcells <- rbind(live, dead)
```

point pattern object 

```{r}
pat_new <- ppp(fullcells[,1], fullcells[,2], c(0,513), c(0,513)) # creating point pattaern object 
marks(pat_new) <- fullcells[,3] # adding infor on live dead - any extra info refered to as mark in spatstat
anyDuplicated(pat_new) # checking for duplicated poitns 
```

extracting how many living and how many dead neighbours in certain radius 

```{r}

countAliveDeadNeighbors <- function(x) {
  alive_count <- sum(x == 1)
  dead_count <- sum(x == 0)
  
  return(c(alive = alive_count, dead = dead_count))
}

neighboursInfo <- t(markstat(pat_new, countAliveDeadNeighbors, R = 15))

```

