---
title: "cell_segmentation_script"
author: "Drew"
date: "2024-02-24"
output: html_document
---

```{r}
library(EBImage)
```

defining functions 

```{r}
plotontop <- function(img, cells, red, xcord, ycord) {
  
  x <- flip(img *2)
  
  par(mar = c(0, 0, 0, 0))  # Set margins to zero
  plot(1, type = "n", xlab = "", ylab = "", xlim = c(0, 1), ylim = c(0, 1), asp =1)
  
  rasterImage(x, 0, 0, 1, dim(x)[2]/dim(x)[1])
  points(x = cells[,xcord] / dim(x)[1], y = cells[,ycord]/ dim(x)[1], col = "blue", pch = 1, cex = 1)
  points(x = red[,xcord] / dim(x)[1], y = red[,ycord] / dim(x)[1], col = "red", pch = 1, cex = 1)
}
```

load original segmentation + image and view 

```{r}
OGgreen <- read.csv("data/1344_48h_Repeat Oct 2020/1344 60min PAA 48hrs 1.green.csv")
OGred <- read.csv("data/1344_48h_Repeat Oct 2020/1344 60min PAA 48hrs 1.red.csv")

img <- readImage("data/PAA1.tif")
display(img, method="browser")
```



defining output name 

```{r}
outputName <- "1_PAA_48_60_1"
```

split colour channels 

```{r}

imgR <- getFrame(img, 1)
imgG <- getFrame(img, 2)

par(mfrow = c(3,1))
plot(imgR)
plot(imgG)
plot(img)

```

offset subtraction - cropped 

```{r}

par(mfrow = c(1,2))
plot(imgG[463:483, 400:420]*4) # examining offset area 
plot(imgR[463:483, 400:420]*4)

osR=mean(imgR[463:483, 400:420]) # taking mean inensity of area 
osG=mean(imgG[463:483, 400:420])

imgRo <- imgR - osR # subtracting mean intensity from full image 
imgGo <- imgG - osG

par(mfrow = c(2,2))
plot(imgR)
plot(imgRo)
plot(imgG)
plot(imgGo)
```

explore image 

```{r}

hist(img)
img

hist(imgRo)
imgRo

hist(imgGo)
imgGo

```

### segmenting live cells 

edge detection

```{r}

k = matrix(-1, nrow = 3, ncol = 3)
k[2, 2] = 8
k

imgGe <- EBImage::filter2(gblur(imgGo,sigma=1.5),k)
par(mfrow = c(2,1))
plot(img)
plot(normalize(imgGe))

```

adaptive thresholding - exploring offset 

```{r}

for (i in seq(from = 0.025, to = 0.5, by = 0.025)) {
  
  x <- thresh(imgGe, w=5, h=5, offset= i)
  
  par(mfrow = c(2,1))
  display(x, method="raster")
  text(x = 20, y = 20, label = paste("offset = ", i), adj = c(0,1), col = "green", cex = 1.5)
  display(img * 2, method="raster")
  text(x = 20, y = 20, label = "Original image", adj = c(0,1), col = "green", cex = 1.5)
  
}

```

selecting best method 

```{r}

for (i in c(0.02, 0.025, 0.03)) {
  
  x <- thresh(imgGe, w=5, h=5, offset= i)
  
  par(mfrow = c(2,1))
  display(x, method="raster")
  text(x = 20, y = 20, label = paste("offset = ", i), adj = c(0,1), col = "green", cex = 1.5)
  display(img * 2, method="raster")
  text(x = 20, y = 20, label = "Original image", adj = c(0,1), col = "green", cex = 1.5)
  
}

imgGth <- thresh(imgGe, w=5, h=5, offset= 0.025)
plot(imgGth)

```

watershed algorithm 

```{r}

nmaskG <- watershed(distmap(imgGth), tolerance = 1, ext = 1)

display(colorLabels(nmaskG), all=TRUE)

par(mfrow = c(2,1))
plot(colorLabels(nmaskG))
plot(img * 2)

```

### extracting red cells 

highlighting red cells - explorign values 

```{r}

for (i in seq(from = 1, to = 3, by = 0.1 )) {

  x <- imgRo - (imgGo / i)

  par(mfrow = c(2,1))
  display(x * 2, method="raster")
  text(x = 20, y = 20, label = paste("ratio = ", i), adj = c(0,1), col = "green", cex = 1.5)
  display(img * 2, method="raster")
  text(x = 20, y = 20, label = "Original image", adj = c(0,1), col = "green", cex = 1.5)

}

imgRb <- imgRo - (imgGo / 3.5)

```

edge detection - red only 

```{r}

k = matrix(-1, nrow = 3, ncol = 3)
k[2, 2] = 8
k

imgRe <- EBImage::filter2(gblur(imgRb,sigma=1.5),k)
par(mfrow = c(2,1))
plot(img)
plot(normalize(imgRe))

```

adaptive thresholding - exploring offset 

```{r}

for (i in seq(from = 0.025, to = 0.4, by = 0.025)) {
  
  x <- thresh(imgRe, w=5, h=5, offset= i)
  
  par(mfrow = c(2,1))
  display(x, method="raster")
  text(x = 20, y = 20, label = paste("offset = ", i), adj = c(0,1), col = "green", cex = 1.5)
  display(img * 2, method="raster")
  text(x = 20, y = 20, label = "Original image", adj = c(0,1), col = "green", cex = 1.5)
  
}

```

adaptive thresholding - selecting best method 

```{r}

imgRth <- thresh(imgRe, w=5, h=5, offset= 0.025)
plot(imgRth)

```

watershed

```{r}

nmaskR <- watershed(distmap(imgRth), tolerance = 1, ext = 1)

#nmaskR <- bwlabel(imgRth)

display(colorLabels(nmaskR), all=TRUE)

par(mfrow = c(2,1))
plot(colorLabels(nmaskR))
plot(img * 2)

```

### inspection of segmentations so far

saving outputs 

```{r}
cellLocR <- as.data.frame(computeFeatures.moment(nmaskR))
cellLocG <- as.data.frame(computeFeatures.moment(nmaskG))
```

plotting segmentation

```{r}
par(mfrow = c(1,1))

plotontop(img, cellLocG, cellLocR, "m.cx", "m.cy")
```

plotting original segmentation 

```{r}
par(mfrow = c(1,1))

plotontop(img, OGgreen, OGred, "XM", "YM")
```

### removing duplicated cells 

```{r}
library(spatstat)
```

manipulating data

```{r}
dead <- cellLocR[,1:2]
dead[,3] <- "0"

living <- cellLocG[,1:2]
living[,3] <- "1"
 
cells <- rbind(dead, living)
colnames(cells) <- c("m.cx", "m.cy", "condition")
```

calculating nearest neighbour and selecting those closet together 

```{r}

NN_which <- nnwhich(cells)
NN_dist <- nndist(cells)

NN_dist_live <- nndist(living)
NN_dist_dead <- nndist(dead)

NN_all <- as.data.frame(cbind(NN_which, NN_dist))

```

examining dist data

```{r}
par(mar = c(10,10,10,10))
par(mfrow = c(1,1))

min(NN_all$NN_dist)
max(NN_all$NN_dist)
hist(NN_all$NN_dist)

min(NN_dist_live)
max(NN_dist_live)
hist(NN_dist_live)

min(NN_dist_dead)
max(NN_dist_dead)
hist(NN_dist_dead)

```

combining cell coords with distance - removing close points 

```{r}

min(NN_dist_live) # checking smallest distance between all cells
min(NN_dist_dead) # checking smallest distance between all cells

cellDists <- cbind(cells, NN_all)

cellLocLive <- as.data.frame(cellDists[cellDists$condition == "1" & cellDists$NN_dist > min(NN_dist_live),])

```

exaimining final segmentation 

```{r}

par(mfrow = c(1,1))

plotontop(img, cellLocLive, cellLocR, "m.cx", "m.cy") # examining my segmentation 

plotontop(img, OGgreen, OGred, "XM", "YM") # comparing to original 

```

### saving output 

```{r}

write.csv(cellLocLive, paste0("results/segmentation/Live_", outputName, ".csv"))
write.csv(cellLocLive, paste0("results/segmentation/Dead_", outputName, ".csv"))

```

