---
title: "Spatial_analysis"
author: "Drew"
date: "2024-02-05"
output: html_document
---

```{r}
#install.packages("spatstat", repos = "https://www.stats.bris.ac.uk/R/")
#install.packages("nortest", repos = "https://www.stats.bris.ac.uk/R/")
```

```{r}
library(spatstat)
```

```{r}

living <- read.csv("results/livingCellsLoc.csv")
dead <- read.csv("results/deadCellsLoc.csv")

```

data prep 

```{r}
living <- living[, c("m.cx", "m.cy")]
living[,"condition"] <- 1

dead <- dead[, c("m.cx", "m.cy")]
dead[,"condition"] <- 0

fullcells <- rbind(living, dead)
```

point pattern object 

```{r}

pat <- ppp(fullcells[,1], fullcells[,2], c(0,1024), c(0,512)) # creating point pattaern object 
marks(pat) <- fullcells[,3] # adding infor on live dead - any extra info refered to as mark in spatstat

patDead <- ppp(dead[,1], dead[,2], c(0,1024), c(0,512)) # creating point pattaern object 

patLive <- ppp(living[,1], living[,2], c(0,1024), c(0,512)) # creating point pattaern object 

```

exploring point pattern objects

```{r}

plot(pat)
anyDuplicated(pat) # checking for duplicated poitns 
summary(pat)

plot(patDead)
anyDuplicated(patDead) 
summary(patDead)

plot(patLive)
anyDuplicated(patLive)
summary(patLive)

```

Spatial randomness test 

```{r}
QAll <- quadratcount(pat, nx = 20, ny = 10)
plot(pat)
plot(QAll, add = TRUE, cex = 1)

quadrat.test(QAll, alternative="two.sided")
quadrat.test(QAll, alternative="regular")
quadrat.test(QAll, alternative="clustered")

QDead <- quadratcount(patDead, nx = 20, ny = 10)
plot(patDead)
plot(QDead, add = TRUE, cex = 1)

quadrat.test(QDead, alternative="two.sided")
quadrat.test(QDead, alternative="regular")
quadrat.test(QDead, alternative="clustered")

QLive <- quadratcount(patLive, nx = 20, ny = 10)
plot(patLive)
plot(QLive, add = TRUE, cex = 1)

quadrat.test(QLive, alternative="two.sided")
quadrat.test(QLive, alternative="regular")
quadrat.test(QLive, alternative="clustered")
```

density tests 

```{r}
denAll <- density(pat) # using default bandwidth 
attr(denAll, "sigma") # shows what bandwidth defaulted to 
plot(denAll, main = "All") #plotting intensities 

denDead <- density(patDead)
attr(denDead, "sigma")
plot(denDead, main = "Dead")

denLive <- density(patLive)
attr(denLive, "sigma")
plot(denLive, main = "Live")

densityVal <- denAll[pat] # extracting intensity values for each point 
```

extracting distance data

```{r}

NNdist <- as.data.frame(nndist(pat, k = 1:3))
NNdist[,"meanDist"] <- rowMeans(NNdist[, 1:3], na.rm = TRUE)
NNdist[,"sumDist"] <- rowSums(NNdist[, 1:3], na.rm = TRUE)

```

extracting how many living and how many dead neighbours in certain radius

```{r}

# defining function to use in the marksat function 
NEcounts <- function(x) {
  alive_count <- sum(x == 1)
  dead_count <- sum(x == 0)
  
  return(c(aliveN = alive_count, deadN = dead_count))
}

neighbourInf <- t(markstat(pat, NEcounts, R = 15))

```

creating data frame of output 

```{r}

cellInf <- cbind(fullcells, densityVal, NNdist, neighbourInf)

deadInf <- cellInf[cellInf$condition == 0 ,]
liveInf <- cellInf[cellInf$condition == 1 ,]

```

exploring outputs 

```{r}

for (i in c("densityVal", "dist.1", "dist.2", "dist.3", "meanDist", "sumDist", "aliveN", "deadN")) {
  
  cat("Variable:", i, "\n")
  
  cat("Mean (Dead):", mean(deadInf[[i]]), "\n")
  cat("Median (Dead):", median(deadInf[[i]]), "\n")

  cat("Mean (Live):", mean(liveInf[[i]]), "\n")
  cat("Median (Live):", median(liveInf[[i]]), "\n")

  # Plot histograms
  hist(deadInf[[i]], main = paste("Histogram (Dead) -", i), xlab = i, col = "lightblue", border = "black")
  hist(liveInf[[i]], main = paste("Histogram (Live) -", i), xlab = i, col = "lightgreen", border = "black")

  # Pause for user input (press Enter to continue)
  cat("Press Enter to continue...")
  invisible(readLines(n = 1))
}


```

saving output 

```{r}

write.csv(deadInf, "results/deadInf.csv")
write.csv(livingInf, "results/livingInf.csv")

```




exploring output

```{r}

library(nortest)

mean(livingInf$sumN)
mean(deadInf$sumN)

mean(livingInf$NNdist)
mean(deadInf$NNdist)

hist(livingInf$sumN)
qqnorm(livingInf$sumN)
qqline(livingInf$sumN)
ad.test(livingInf$sumN)

hist(deadInf$sumN)
qqnorm(deadInf$sumN)
qqline(deadInf$sumN)
ad.test(deadInf$sumN)
shapiro.test(deadInf$sumN)

hist(livingInf$NNdist)
qqnorm(livingInf$NNdist)
qqline(livingInf$NNdist)
ad.test(livingInf$NNdist)

hist(deadInf$NNdist)
qqnorm(deadInf$NNdist)
qqline(deadInf$NNdist)
ad.test(deadInf$NNdist)
shapiro.test(deadInf$NNdist)

```

saving outputs 

```{r}

write.csv(deadInf, "results/deadInf.csv")
write.csv(livingInf, "results/livingInf.csv")

```




extracting distance to closet neighbour 

```{r}
NNdist <- nndist(pat, k = 1:3)
k3dists <- cbind(fullcells, NNdist) 
str(k3dists)
k3dists[,"meanDist"] <- rowMeans(k3dists[, 4:6], na.rm = TRUE)
k3dists[,"sumDist"] <- rowSums(k3dists[, 4:6], na.rm = TRUE)

#livetest <- k3dists[k3dists$condition == "1" ,]
#deadtest <- k3dists[k3dists$condition == "0" ,]

#mean(livetest$sumDist)
#mean(deadtest$sumDist)

#mean(livetest$meanDist)
#mean(deadtest$meanDist)
```








```{r}
#img <- readImage("results/img.tiff")
red <- read.csv("results/redcells.csv")
green <- read.csv("results/greencells.csv")
```

```{r}
red <- red[,2:3]
red[,3] <- "dead"

green <- green[,2:3]
green[,3] <- "live"
 
allcells <- rbind(red, green)
colnames(allcells) <- c("x", "y", "condition")

min(allcells$x)
min(allcells$y)
max(allcells$x)
max(allcells$y)
```

creating spatstat point pattern object 

```{r}
pat <- ppp(allcells[,1], allcells[,2], c(0,513), c(0,513)) # creating point pattaern object 
marks(pat) <- allcells[,3] # adding infor on live dead - any extra info refered to as mark in spatstat
anyDuplicated(pat) # checking for duplicated poitns 
```

general descriptive statistics

```{r}
# global_density <- length(bike_theft_2019$NAME)/sum(st_area(london_ward_shp))
```



```{r}
plot(myPattern)
summary(myPattern)
plot(Kest(myPattern))
plot(density(myPattern))
plot(smooth(myPattern))
```

```{r}
radius <- 0.1
n_neighbors <- nncross(myPattern, k = 0, r = radius, by = "marks")
```

```{r}
radius <- 2  # replace with your desired radius
neighbors <- nncross(allcells$x, allcells$y, r = radius)
```

```{r}
W <- square(1)
set.seed(42)
Y <- runifpoint(2000) # Random points in the unit square
plot(Y, main = "Random points in unit square")
```

```{r}
xy <- gridcenters(W, 513, 513) # Grid of points in the unit square
X <- ppp(xy$x, xy$y, window = W, check = FALSE, checkdup = FALSE)
plot(X)
```

```{r}
test <- crosspaircounts(myPattern, X, r = 0.5)
```

could also exract info such as of the 10 neareast neighbours, how many are dead and how many are alive 