---
title: "cell_segmentation_script"
author: "Drew"
date: "2024-02-24"
output: html_document
---

```{r}
library(spatstat)
library(ggplot2)
library(gridExtra)
library(EBImage)
```

defining functions 

```{r}
plotontop <- function(img, cells, red, xcord, ycord, title) {
  
  x <- flip(img *2)
  
  par(mar = c(0, 0, 0, 0))  # Set margins to zero
  plot(1, type = "n", xlab = "", ylab = "", xlim = c(0, 1), ylim = c(0, 1), asp =1, main = paste(title))
  
  rasterImage(x, 0, 0, 1, dim(x)[2]/dim(x)[1])
  points(x = cells[,xcord] / dim(x)[1], y = cells[,ycord]/ dim(x)[1], col = "blue", pch = 1, cex = 0.4)
  points(x = red[,xcord] / dim(x)[1], y = red[,ycord] / dim(x)[1], col = "red", pch = 1, cex = 0.4)
}
```

defining variable names 

```{r}
inputDir <- "data/1344_48h_Repeat Oct 2020"
inputName <- "1344 20min H2O2 48hrs 3"
outputName <- "1_48_H2O2_20_3"
```

load original segmentation + image and view 

```{r}
OGgreen <- read.csv(paste0(inputDir, "/", inputName, ".green.csv"))
OGred <- read.csv(paste0(inputDir, "/", inputName, ".red.csv"))

img <- readImage(paste0(inputDir, "/", inputName, ".tif"))
display(flip(img *2), method="browser")
display(img *2, method="browser")
```

defining script variables and which parts to run

```{r}
thGvalue <- 0.028
thRvalue <- 0.045
rRatioValue <- 1.5
gblurVal <- 1.5

thGloop <- "n"
thRloop <- "n"
rRatioLoop <- "n"
offsetPos <- "n"
#redOrGreenDominant <- "r"


saveOut <- "n"

offsetRegion <- "NA"
```

split colour channels 

```{r}

imgR <- getFrame(img, 1)
imgG <- getFrame(img, 2)

par(mfrow = c(3,1))
plot(imgR)
plot(imgG)
plot(img)

```

offset subtraction - cropped 

```{r}

if(offsetPos == "y") {
  
  par(mfrow = c(1,2))
  plot(imgG[654:664, 319:329]*4)
  plot(imgR[654:664, 319:329]*4)

  osR=mean(imgR[654:664, 319:329])
  osG=mean(imgG[654:664, 319:329])

  imgRo <- normalize(imgR - osR, ft=c(0,1)) # subtracting mean intensity from full image 
  imgGo <- normalize(imgG - osG, ft=c(0,1))

  par(mfrow = c(2,2))
  plot(imgR)
  plot(imgRo)
  plot(imgG)
  plot(imgGo)

} else {
  
  cat("no offset possible")
  
  imgRo <- normalize(imgR, ft=c(0,1))
  imgGo <- normalize(imgG, ft=c(0,1))
  
}
```

explore image 

```{r}

par(mar = c(5,5,5,5))
par(mfrow = c(2,2))

hist(imgRo)
imgRo

hist(imgGo)
imgGo

hist(img)
img

```

### segmenting live cells 

edge detection

```{r}

k = matrix(-1, nrow = 3, ncol = 3)
k[2, 2] = 8
k

imgGe <- EBImage::filter2(gblur(imgGo,sigma=gblurVal),k)
par(mfrow = c(2,1))
plot(img)
plot(normalize(imgGe))

```

adaptive thresholding - exploring offset 

```{r}

if(thGloop == "y") {
  
  for (i in seq(from = 0.005, to = 0.05, by = 0.005)) {
  
    x <- thresh(imgGe, w=5, h=5, offset= i)
  
    par(mfrow = c(2,1))
    display(x, method="raster")
    text(x = 20, y = 20, label = paste("offset = ", i), adj = c(0,1), col = "green", cex = 1.5)
    display(img * 2, method="raster")
    text(x = 20, y = 20, label = "Original image", adj = c(0,1), col = "green", cex = 1.5)
  
  }
} else {
  cat("green threshold loop not run")
}

```

selecting best method 

```{r}

imgGth <- thresh(imgGe, w=5, h=5, offset = thGvalue)
plot(imgGth)

```

watershed algorithm 

```{r}

nmaskG <- watershed(EBImage::distmap(imgGth), tolerance = 1, ext = 1)

#display(colorLabels(nmaskG), all=TRUE)

par(mfrow = c(2,1))
plot(colorLabels(nmaskG))
plot(img * 2)

```

### extracting red cells 

highlighting red cells - explorign values 

```{r}

if(rRatioLoop == "y") {
  
  for (i in seq(from = 1, to = 3, by = 0.1 )) {

    x <- imgRo - (imgGo / i)

    par(mfrow = c(2,1))
    display(x * 2, method="raster")
    text(x = 20, y = 20, label = paste("ratio = ", i), adj = c(0,1), col = "green", cex = 1.5)
    display(img * 2, method="raster")
    text(x = 20, y = 20, label = "Original image", adj = c(0,1), col = "green", cex = 1.5)

  }
} else {
  cat("red ratio loop not run")
}

imgRb <- imgRo - (imgGo / rRatioValue)
plot(imgRb)

```

edge detection - red only 

```{r}

k = matrix(-1, nrow = 3, ncol = 3)
k[2, 2] = 8
k

imgRe <- EBImage::filter2(gblur(imgRb,sigma=1.5),k)
par(mfrow = c(2,1))
#plot(img)
#plot(normalize(imgRe))

```

adaptive thresholding - exploring offset 

```{r}

if(thRloop == "y") {
  
  for (i in seq(from = 0.005, to = 0.05, by = 0.005)) {
  
    x <- thresh(imgRe, w=5, h=5, offset= i)
  
    par(mfrow = c(2,1))
    display(x, method="raster")
    text(x = 20, y = 20, label = paste("offset = ", i), adj = c(0,1), col = "green", cex = 1.5)
    display(img * 2, method="raster")
    text(x = 20, y = 20, label = "Original image", adj = c(0,1), col = "green", cex = 1.5)
  
  }
} else {
  cat("red threshold loop not run")
}

```

adaptive thresholding - selecting best method 

```{r}

imgRth <- thresh(imgRe, w=5, h=5, offset= thRvalue)
plot(imgRth)

```

watershed

```{r}

nmaskR <- watershed(EBImage::distmap(imgRth), tolerance = 1, ext = 1)

#nmaskR <- bwlabel(imgRth)

#display(colorLabels(nmaskR), all=TRUE)

par(mfrow = c(2,1))
plot(colorLabels(nmaskR))
plot(img * 2)

```

extracting intenstiy values 

```{r}

IntValDeadR <- as.data.frame(computeFeatures.basic(nmaskR, imgRo)[,1:2])
IntValDeadG <- as.data.frame(computeFeatures.basic(nmaskR, imgGo)[,1:2])

IntValLiveR <- as.data.frame(computeFeatures.basic(nmaskG, imgRo)[,1:2])
IntValLiveG <- as.data.frame(computeFeatures.basic(nmaskG, imgGo)[,1:2])

```

organising outputs

```{r}
cellLocR <- as.data.frame(cbind(computeFeatures.moment(nmaskR)[,1:2], IntValDeadR, IntValDeadG))
cellLocR$condition <- "0"
colnames(cellLocR) <- c("m.cx", "m.cy","redMeanInts", "redSDInts", "greenMeanInts", "greenSDInts", "condition")

cellLocG <- as.data.frame(cbind(computeFeatures.moment(nmaskG)[,1:2], IntValLiveR, IntValLiveG))
cellLocG$condition <- "1"
colnames(cellLocG) <- c("m.cx", "m.cy","redMeanInts", "redSDInts", "greenMeanInts", "greenSDInts", "condition")

cellLocFull <- rbind(cellLocR, cellLocG)
```

exaimining first segmentation

```{r}

#par(mfrow = c(1,1))

#plotontop(img, cellLocG, cellLocR, "m.cx", "m.cy", "first segmentation") # examining my segmentation 

#plotontop(img, OGgreen, OGred, "XM", "YM", "original segmentation") # comparing to original 

```


calculating nearest neighbour and selecting those closet together 

```{r}

cellLocFull$NN_dist <- nndist(cellLocFull[, 1:2])

cellLocR$NN_dist <- nndist(cellLocR[, 1:2])
cellLocG$NN_dist <- nndist(cellLocG[, 1:2])

```

examining dist data

```{r}
par(mar = c(4,4,4,4))
par(mfrow = c(2,2))

min(cellLocFull$NN_dist)
max(cellLocFull$NN_dist)
hist(cellLocFull$NN_dist)

min(cellLocR$NN_dist)
max(cellLocR$NN_dist)
hist(cellLocR$NN_dist)

min(cellLocG$NN_dist)
max(cellLocG$NN_dist)
hist(cellLocG$NN_dist)

```

combining cell coords with distance - removing close points 

```{r}

min_NN_dist <- min(min(cellLocG$NN_dist), min(cellLocR$NN_dist))

cellLocGadj <- as.data.frame(cellLocFull[cellLocFull$condition == "1" & cellLocFull$NN_dist > min_NN_dist,])

cellLocFullAdj <- rbind(cellLocGadj,cellLocR)

```

exaimining second segmentation

```{r}

#par(mfrow = c(1,1))

#plotontop(img, cellLocGadj, cellLocR, "m.cx", "m.cy", "second segmentation") # examining my segmentation 

#plotontop(img, OGgreen, OGred, "XM", "YM", "original segmentation") # comparing to original 

```

setting threshold cutoff 

```{r}

cellLocFullAdj09 <- cellLocFullAdj
cellLocFullAdj1 <- cellLocFullAdj

deadSet09 <- which(cellLocFullAdj09$redMeanInts / cellLocFullAdj09$greenMeanInts > 0.9)
liveSet09 <- which(cellLocFullAdj09$redMeanInts / cellLocFullAdj09$greenMeanInts < 0.9)

cellLocFullAdj09$condition[deadSet09] <- "0" 
cellLocFullAdj09$condition[liveSet09] <- "1" 

deadSet1 <- which(cellLocFullAdj1$redMeanInts > cellLocFullAdj1$greenMeanInts)
liveSet1 <- which(cellLocFullAdj1$greenMeanInts > cellLocFullAdj1$redMeanInts)

cellLocFullAdj1$condition[deadSet1] <- "0" 
cellLocFullAdj1$condition[liveSet1] <- "1" 

```

viewing adjusted segmentation 

```{r}

par(mfrow = c(1,1))

plotontop(img, OGgreen, OGred, "XM", "YM", "original segmentation") # comparing to original 

plotontop(img, cellLocFullAdj[cellLocFullAdj$condition == "1" ,], 
          cellLocFullAdj[cellLocFullAdj$condition == "0" ,], "m.cx", "m.cy", "standard") # examining my segmentation 

plotontop(img, cellLocFullAdj09[cellLocFullAdj09$condition == "1" ,], 
          cellLocFullAdj09[cellLocFullAdj09$condition == "0" ,], "m.cx", "m.cy", "0.9 threshold") # examining my segmentation 

plotontop(img, cellLocFullAdj1[cellLocFullAdj1$condition == "1" ,], 
          cellLocFullAdj1[cellLocFullAdj1$condition == "0" ,], "m.cx", "m.cy", "1 threshold") # examining my segmentation 


p1 <- ggplot(cellLocFullAdj, mapping = aes(x = redMeanInts, y = greenMeanInts, colour = condition)) +
  geom_point() +
  xlim(c(0,1)) +
  ylim(c(0,1)) +
  ggtitle("original segmentation")

p2 <- ggplot(cellLocFullAdj09, mapping = aes(x = redMeanInts, y = greenMeanInts, colour = condition)) +
  geom_point() +
  xlim(c(0,1)) +
  ylim(c(0,1)) +
  ggtitle("0.9 threshold")

p3 <- ggplot(cellLocFullAdj1, mapping = aes(x = redMeanInts, y = greenMeanInts, colour = condition)) +
  geom_point() +
  xlim(c(0,1)) +
  ylim(c(0,1)) +
  ggtitle("1:1 threshold")

grid.arrange(p1, p2, p3, ncol = 3)

```

### dataframe for segmentation values 

```{r}
varNames <- c("thGvalue", "thRvalue", "rRatioValue", "offsetRegion", "sumR", "sumG")
varVal <- c(thGvalue, thRvalue, rRatioValue, offsetRegion, sum(imgR), sum(imgG))

segVals <- data.frame(varNames = varNames, varVal = varVal)
```

### saving output 

```{r}
if(saveOut == "y") {
  
  write.csv(cellLocFullAdj, paste0("outputs/segmentation/", outputName, ".csv"))
  write.table(segVals, paste0("outputs/segmentation/", outputName, "_values.txt"), sep = "\t")
  
  write.csv(cellLocFullAdj09, paste0("outputs/segmentation09/", outputName, ".csv"))
  write.table(segVals, paste0("outputs/segmentation09/", outputName, "_values.txt"), sep = "\t")
  
  write.csv(cellLocFullAdj1, paste0("outputs/segmentation1/", outputName, ".csv"))
  write.table(segVals, paste0("outputs/segmentation1/", outputName, "_values.txt"), sep = "\t")
  
} else {
  cat("not saved")
}
```
