---
title: "Untitled"
author: "Drew"
date: "2024-03-28"
output: html_document
---

```{r}
library(spatstat)
library(ggplot2)
library(gridExtra)
library(EBImage)
```

```{r}
plotontop <- function(img, cells, red, xcord, ycord, title) {
  
  x <- flip(img *3)
  
  par(mar = c(0, 0, 0, 0))  # Set margins to zero
  plot(1, type = "n", xlab = "", ylab = "", xlim = c(0, 1), ylim = c(0, 1), asp =1, main = paste(title))
  
  rasterImage(x, 0, 0, 1, dim(x)[2]/dim(x)[1])
  points(x = cells[,xcord] / dim(x)[1], y = cells[,ycord]/ dim(x)[1], col = "blue", pch = 1, cex = 1)
  points(x = red[,xcord] / dim(x)[1], y = red[,ycord] / dim(x)[1], col = "red", pch = 1, cex = 1)
}
```

defining variable names 

```{r}
inputDir <- "data/1344_48h_Repeat Oct 2020"
inputName <- "1344 20min H2O2 48hrs 1"
outputName <- "1_48_H2O2_20_1"
```

load original segmentation + image and view 

```{r}
OGgreen <- read.csv(paste0(inputDir, "/", inputName, ".green.csv"))
OGred <- read.csv(paste0(inputDir, "/", inputName, ".red.csv"))

imgRaw <- readImage(paste0(inputDir, "/", inputName, ".tif"))
display(flip(imgRaw *2), method="browser")
display(imgRaw *2, method="browser")
```

setting segmentation variables 

```{r}
threshVal <- 0.02
gblurVal <- 1.5
#offsetRegion <- list(r = 654:664, c = 319:329)
cropRegion <- list(r = 256:984, c = 1:512)
offsetRegion <- "NA"
#cropRegion <- "NA"
redAdj <- 0.5
greenAdj <- 1
comments <- "blury image, very low red intensity, issues with combinging spectrums"

threshLoop <- "y"
offsetPos <- "n"
cropReq <- "y"

saveOut <- "n"
```

cropping image 

```{r}
if(cropReq == "y") {
  
  img <- imgRaw[cropRegion$r, cropRegion$c, 1:3]
  
  par(mfrow = c(1,1))
  plot(img)


} else {
  
  cat("no cropping required")
  
  img <- imgRaw
  
}
```


splitting and combining colour channels 

```{r}
imgR <- getFrame(img, 1)
imgG <- getFrame(img, 2)

imgC <- normalize((imgR * redAdj) + (imgG * greenAdj), ft=c(0,1))

par(mfrow=c(3,1))
plot(imgR)
plot(imgC)
plot(imgG)
```

offset subtraction 

```{r}
if(offsetPos == "y") {
  
  par(mfrow = c(1,2))
  plot(imgC[offsetRegion$r, offsetRegion$c]*4)

  os=mean(imgC[offsetRegion$r, offsetRegion$c])

  imgCo <- normalize(imgC - os, ft=c(0,1)) # subtracting mean intensity from full image 

  par(mfrow = c(1,2))
  plot(imgC)
  plot(imgCo)


} else {
  
  cat("no offset possible")
  
  imgCo <- normalize(imgC, ft=c(0,1))
  
}
```

edge detection

```{r}
k = matrix(-1, nrow = 3, ncol = 3)
k[2, 2] = 8
k

imgE <- EBImage::filter2(gblur(imgCo,sigma=gblurVal),k)
par(mfrow = c(2,1))
plot(img)
plot(normalize(imgE))
```

selecting best method 

```{r}
if(threshLoop == "y") {
  
  for (i in seq(from = 0.005, to = 0.05, by = 0.005)) {
  
    x <- thresh(imgE, w=5, h=5, offset= i)
  
    par(mfrow = c(2,1))
    display(x, method="raster")
    text(x = 20, y = 20, label = paste("offset = ", i), adj = c(0,1), col = "green", cex = 1.5)
    display(img * 2, method="raster")
    text(x = 20, y = 20, label = "Original image", adj = c(0,1), col = "green", cex = 1.5)
  
  }
} else {
  cat("green threshold loop not run")
}

imgTh <- thresh(imgE, w=5, h=5, offset = threshVal)
plot(img)
plot(imgTh)
```

eroding + dilation romeove single pixels 

```{r}
#kern <- makeBrush(3, shape = "box")
#logo_erode= erode(imgTh, kern)
#logo_dilate = dilate(logo_erode, kern)

#plot(logo_dilate)
#plot(imgTh)
```


watershed algorithm 

```{r}
nmask <- watershed(EBImage::distmap(imgTh), tolerance = 1, ext = 1)

par(mfrow = c(2,1))
plot(colorLabels(nmask), all = T)
plot(img * 2)
```

extracting intensity vals and cell coordinates 

```{r}
IntValR <- as.data.frame(computeFeatures.basic(nmask, imgR)[,1:2])
IntValG <- as.data.frame(computeFeatures.basic(nmask, imgG)[,1:2])

cellLoc <- as.data.frame(cbind(computeFeatures.moment(nmask)[,1:2], IntValR, IntValG))
colnames(cellLoc) <- c("m.cx", "m.cy","redMeanInts", "redSDInts", "greenMeanInts", "greenSDInts")
cellLoc$condition <- "null"

deadSet <- which(cellLoc$redMeanInts > cellLoc$greenMeanInts)
liveSet <- which(cellLoc$greenMeanInts > cellLoc$redMeanInts)

cellLoc$condition[deadSet] <- "0" 
cellLoc$condition[liveSet] <- "1" 
```

plotting original and new segmentation 

```{r}
par(mfrow = c(1,1))

ggplot(cellLoc, mapping = aes(x = redMeanInts, y = greenMeanInts, colour = condition)) +
  geom_point() +
  xlim(c(0,1)) +
  ylim(c(0,1)) +
  ggtitle("live dead distribution") + 
  theme(aspect.ratio=1)

plotontop(imgRaw, OGgreen, OGred, "XM", "YM", "original segmentation") # comparing to original 

plotontop(img, cellLoc[cellLoc$condition == "1" ,], 
          cellLoc[cellLoc$condition == "0" ,], "m.cx", "m.cy", "new segmentation") # examining my segmentation 
```

dataframe for segmentation values 

```{r}
varNames <- c("threshVal", "gblurVal", "offsetRegion", "cropRegion", "sumR", "sumG", 
            "meanG", "meanR", "redAdj", "greenAdj", "comments")

varVal <- c(threshVal, gblurVal, paste(offsetRegion, collapse = " "), 
            paste(cropRegion, collapse = " "), sum(imgR), sum(imgG), 
            mean(cellLoc$greenMeanInts), mean(cellLoc$redMeanInts), redAdj, greenAdj, comments)

segVals <- data.frame(varNames = varNames, varVal = varVal)
```

### saving output 

```{r}
if(saveOut == "y") {
  
  write.csv(cellLoc, paste0("results_good/segmentation/", outputName, ".csv"))
  write.table(segVals, paste0("results_good/segmentation/", outputName, "_values.txt"), sep = "\t")
  
} else {
  cat("not saved")
}
```
