---
title: "first_experiments"
author: "Drew"
date: "2024-01-12"
output: html_document
---

```{r}
BiocManager::install("EBImage")
```

```{r}
library(EBImage)
```

loading image in

```{r}
img <- readImage("data/trial2.tif")

writeImage(img, "test.jpeg")
```

viewing image 

```{r}
display(img, method="browser") # using browser method runs java script through web brower to view object 
display(img, method="raster") # uses r's default method allows combo with other plotting functionality to add labels etc 
```

expolration of image 

```{r}
hist(img) # hist look at colour frequency distrbution 
img # display general stats about image 
```

```{r}
imgRt <- getFrame(img, 1)

imgGbase <- channel(img, "green") # extracting green spectrum 
imgRbase <- channel(img, "red")
imgBbase <- channel(img, "blue")

display(imgBbase, method="browser") # dispalying extracted spectrum - greyscale as only 1 colour channel 
display(imgRbase, method="browser")
display(imgGbase, method="browser")

imgG <- rgbImage(red = 0 * imgRbase, green = imgGbase, blue = 0 * imgBbase) # coonverting image back to colour image
imgR <- rgbImage(red = imgRbase, green = 0 * imgGbase, blue = 0 * imgBbase)
imgB <- rgbImage(red = 0 * imgRbase, green = 0 * imgGbase, blue = imgBbase)

display(imgB, method="browser") # displaying col image 
display(imgR, method="browser")
display(imgG, method="browser")

display(img, method="browser")

```

investigating extracted colour channels 

```{r}
hist(imgGbase)
imgGbase

hist(imgRbase)
imgRbase
```

noise filtering 

```{r}

Gmfilt <- medianFilter(imgGbase, 3)

display(imgGbase, method="browser")
display(Gmfilt, method="browser")

Ggfilt <- gblur(imgGbase, sigma = 5)

display(imgGbase, method="browser")
display(Ggfilt, method="browser")




nfilt <- medianFilter(img, 3)

display(nfilt, method="browser")
display(img, method="browser")

hist(img)

```

```{r}
k = matrix(-1, nrow = 3, ncol = 3)
k[2, 2] = 8
k

Geimg=EBImage::filter2(gblur(imgGbase,sigma=1.5),k)
display(Geimg, method="browser")

Reimg=EBImage::filter2(gblur(imgRbase,sigma=1.5),k)
display(Reimg, method="browser")
```


thresholding

```{r}
Gth <- otsu(imgGbase)
Gthresh <- imgGbase>0.5
display(Gthresh, method="browser")

Rth <- otsu(imgRbase)
Rthresh <- imgRbase>0.5
display(Rthresh, method="browser")

```

this seems like a nice method as it preserves a lot of spatial information 

```{r}

Gat_img=EBImage::thresh(CimgGbase,h = 5,w = 5, offset = 0.15)

display(Gat_img, method="browser")
display(imgGbase, method="browser")

Rat_img=EBImage::thresh(CimgRbase,h = 5,w = 5, offset = 0.15)

display(Rat_img, method="browser")
display(imgRbase, method="browser")

display(img, method="browser")
```

subtracting green channel from red channel - works pretty good, want to try using just each colour channel and massivley raising contrast? might be good to try 

```{r}
test <- imgRbase - imgGbase 


display(test, method="browser")

#Gth <- otsu(imgGbase)
Gthresh <- test>0.08
display(Gthresh, method="browser")

display(img, method="browser")
```


```{r}
Cimg <- img * 2
display(Cimg, method="browser")

CimgGbase <- channel(Cimg, "green") # extracting green spectrum 
CimgRbase <- channel(Cimg, "red")
CimgBbase <- channel(Cimg, "blue")

display(CimgBbase, method="browser") # dispalying extracted spectrum - greyscale as only 1 colour channel 
display(CimgRbase, method="browser")
display(CimgGbase, method="browser")

CimgG <- rgbImage(red = 0 * CimgRbase, green = CimgGbase, blue = 0 * CimgBbase) # coonverting image back to colour image
CimgR <- rgbImage(red = CimgRbase, green = 0 * CimgGbase, blue = 0 * CimgBbase)
CimgB <- rgbImage(red = 0 * CimgRbase, green = 0 * CimgGbase, blue = CimgBbase)

display(CimgB, method="browser") # displaying col image 
display(CimgR, method="browser")
display(CimgG, method="browser")

display(Cimg, method="browser")

hist(CimgGbase)
CimgGbase

hist(CimgRbase)
CimgRbase

test <- CimgRbase - CimgGbase 


display(test, method="browser")

#Gth <- otsu(imgGbase)
Gthresh <- test>0.2
display(Gthresh, method="browser")

display(img, method="browser")

```

```{r}

```

for apply either dead or alive could use some if funcition in a loop running over all points in both immage arrays to colour the cell as red or green 
